On The Impact of Atoms of Confusion in JavaScript Code

### 2021-01-27 3_Method.tex
revisao adicional na secao de study settings

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lay the foundations for the implementation of libraries that can automatically transform code into its cleaner versions---though we postpone these results to a future research work. To answer these research questions we conduct a mixed-methods study, including {\color{red}two surveys}, a set of interviews, and an effort of mining JavaScript software repositories.

\subsection{First Study: Survey-I}

In the first study we explore the impact of atoms of confusion on
understanding JavaScript code, by answering two questions: (a) \emph{Do code snippets that contain atoms of confusion produce a higher misunderstanding rate when programmers try to predict their outcome?} and (b) \emph{Do code snippets with atoms of confusion require programmers to take longer to predict their output?}
Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs. Appendix~\ref{sec:appendix-atoms} summarizes all atoms we consider in our research.

\subsubsection{Survey design}

@@

  \begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.50]{images/latin-square.pdf}
      \caption{Latin square design. Each ``square'' corresponds to 
      a replica in our study. Each replica comprises two students (square rows) 
      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
      treatments (atom or non-atom code) to the cells of the squares.} 
      \label{fig:latinsquare}
  \end{figure}

Having selected 10 atoms in our first study, we wrote small pieces of code that contained each atom. We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets. In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 listings, wherein each subset contained 5 blocks that contained atoms of confusion, whilst the remaining 5 had the atoms removed. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
We measured answer correctness and the total time each participant needed to answer the survey's questions.

\subsubsection{Survey Instrument}

@@

We organized the survey in {\color{red}three} sections. The first section aims to characterize the subjects, asking for their age, education level, and programming experience. We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes. In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it. We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The next section of the survey presented a sequence of 10 questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

@@

\subsubsection{Survey audience}

@@

\subsection{Second Study: Interviews with Practitioners}

To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion. We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as confusing. In this section we details the protocol we followed to conduct the interviews and analyze the results.

@@

\subsubsection{Participants Selection} We invited the participants of the interviews using a snow balling technique. That is, starting from our network of contacts, we invited an initial set of candidates to take part of our survey. From this initial list, we asked for indication of additional candidates. Our main criterion of selection was that all participants should have been working professionally with JavaScript. We invited a total of 17 developers, and 15 of these agreed to participate. We conducted the interviews during a period of two weeks.

\subsubsection{Interview Process} We conducted semi-structured interviews using a web conferencing software. All the interviews were recorded with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes, and the longest one 43.06 minutes. The main goal of the interviews was to have developers assess in real time whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed in readability, and, if so, which version they regarded as the easier one to understand. The interviews were conducted by two of the researchers, and a third one listened to all the recordings to cross-validate the collected data.

@@

\subsubsection{Interview Analysis}

{\color{red}Nós realizamos a análise das entrevistas em pares. Inicialmente um dos participantes de todas as entrevistas transcreveu os resultados. Posteriormente, um dos autores do trabalho ouviu os áudios das entrevistas e transcreveu os resultados.}

\textcolor{red}{Adriano teria algo mais para adicionar aqui na  análise das entrevistas?}

\todo[inline]{adriano:estou escrevendo uma tabela em que cada linha e um dos atomos, e colunas representam \% preferencia com atomo, \% preferencia sem atomo, \% indiferente. Mas acredito que esta tabela deve estar na section results, certo? O resto das informacoes deste paragrafo eu citei acima no Interview Process, entao acho que podemos abolir esta section. -- Edna: Concordo contigo Adriano}

\subsection{Third Study: Additional Survey and Interviews}

\rb{acho que aqui podemos incluir o segundo survey e a segunda
  rodada das entrevistas, explorando os \'{a}tomos adicionais}

\subsection{Fourth Study: Mining Software Repositories}

To understand how often atoms of confusion appear in real settings, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

@@

In the third step we mined atoms of confusion from the repositories in our curate dataset, using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as Java and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our
analysis (and also the reproduction of this study). In the final step we carry
out basic descriptive statistics, in order to understand how often the
atoms of confusion appear in practice.

>>>>

Answering these research questions has several implications. For instance, we can either generalize or refute the perceptions about atoms of confusion already discussed in the literature (goal of research questions (RQ1) and (RQ2)). In addition, answers to the third and fourth questions allow us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lay the foundations for the implementation of tools that can automatically transform code into its cleaner versions---though we postpone these results to a future research work. To answer these research questions we conduct a mixed-methods study, including a survey, a set of interviews, and an effort of mining JavaScript software repositories.

\subsection{Survey Study Settings}

In the first study we explore the impact of atom candidates on
understanding JavaScript code, by answering two questions:
(a) \emph{Do code snippets that contain atom candidates
produce a higher misunderstanding rate when programmers
try to predict their outcome?}
and (b) \emph{Do code snippets with atom candidates require
programmers to take longer to predict their output?}
Since JavaScript and C have some constructs in common (Section~\ref{sec:aoc}), we first selected a set of atoms of confusion for the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} whose idioms also exist in JavaScript programs. Our supplementary material
summarizes all atoms we consider in our research.

\subsubsection*{Survey design}

@@

  \begin{figure}[htb!]
      \noindent
      \centering
      \includegraphics[scale=.50]{images/latin-square.pdf}
      \caption{Latin square design. Each ``square'' corresponds to 
      a replica in our study. Each replica comprises two participants (square rows) 
      and two sets of code snippets (CS-1 and CS-2). We randomly apply the 
      treatments (atom or non-atom code) to the cells of the squares.} 
      \label{fig:latinsquare}
  \end{figure}

Having selected 10 atoms in our first study, we wrote small pieces of code that contained each atom. We also wrote their equivalent snippet without the confusing idioms and constructs, leading to a total of 20 code snippets. In order to reduce the cognitive effort, we decided that each subject would be asked to predict the output of a subset of 10 code snippets, wherein each subset contained 5 blocks of code that contained atoms of confusion, whilst the remaining 5 had the atoms removed. The order in which the questions were presented was randomized. By doing this, we were seeking to minimize the chances of subjects being aware that the current listing they were analyzing contained (or not) atoms of confusion.
That is, each respondent of the survey should indicate what would be the outcomes of the code snippets, some of them having atoms of confusion (while other code snippets did not). 
We measured answer correctness and the total time each participant needed to answer the survey's questions.

\subsubsection*{Survey Instrument}

@@

We organized the survey in three sections. The first section aims to characterize the subjects, asking for their age, education level, and programming experience. We also included a check button, whose checking meant users agreed that all collected data would be used solely for research purposes. In the second section, we presented to the participants a small set of instructions, where we explained how the survey worked and asked them to dedicate their attention to it. We stressed to participants the importance of not using any aids during the survey, such as online or console interpreters. For each question page, we kept track of whether or not the subjects switched windows.

The last section of the survey presented a sequence of 10 questions, each containing a code snippet. For each question, there was a text box where the answer should be written. There was also an ``I do not know'' button, which, when clicked, led the subject to the next question. In our setting, ``I do not know'' was treated as a wrong answer. The code snippets were presented as images copied from a text editor, so as to demotivate respondents from resorting to external resources by copying and pasting the code into an interpreter. Upon submitting their answer for a particular question, the subject was automatically led to a similar page, containing another snippet.

@@

\subsubsection*{Survey audience}

@@

\subsection{Interview Study Settings}

To complement our initial survey, we performed semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding code snippets containing atoms of confusion. We also asked each participant if they knew of any other JavaScript-specific construct that they regarded as being likely to make the code hard to understand. In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

@@

\subsubsection*{Participants Selection} We invited the participants of the interviews using a snowballing technique. That is, starting from our network of contacts, we invited an initial set of candidates to take part in our survey. From this initial list, we asked for an indication of additional candidates. Our main selection criterion was that all participants should have been working with JavaScript in their daily professional activities. We invited a total of 17 developers, and 15 of them agreed to participate.
%We conducted the interviews for a period of two weeks.

\subsubsection*{Interview Process} We conducted semi-structured interviews using a web conferencing software. We recorded all the interviews with the consent of the participants. On average, the interviews lasted 26.29 minutes, with the shortest one lasting 14.59 minutes and the longest one 43.06 minutes. The main goal of the interviews was to have developers assess whether two versions of code (one with a single atom and another one without the atom) with the same behavior differed concerning program comprehension, and, if so, which version they\
 regarded as the easier one to understand. Two of the researchers conducted the interviews, and a third one listened to all the recordings to cross-validate the collected data.

@@

\subsubsection*{Interview Analysis}

We first transcribed each interview and then examined the broad distribution of the answers. Our goal was to build an initial understanding of the participants' perceptions with regards to the challenges to understand JavaScript code in general and JavaScript code with atoms of confusion. We next followed with an open-coding	procedure, highlighting	the main themes and quoting the participant answers, which we show the most important ones in
Section~\ref{sec:interview-results}. Finally, we consolidate the feedback of the participants with respect to the confusion that our set of atom candidates might introduce in JavaScript code.

%\subsection{Third Study: Additional Survey and Interviews}

% \rb{acho que aqui podemos incluir o segundo survey e a segunda
%  rodada das entrevistas, explorando os \'{a}tomos adicionais}

\subsection{Mining JavaScript Software Repositories Settings}

To understand how often atoms of confusion appear in open source projects, and thus answer our fourth research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available on  GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

@@

We mined the occurrence of atom candidates from the repositories in our curate dataset using source code queries that we had written using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our
analysis (and also the reproduction of this study). Finally, we carried
out basic descriptive statistics, in order to understand how often the
atoms of confusion appear in practice.

### 2021-01-27 main.tex
revisao adicional na secao de study settings

\begin{abstract}
  {\bf Context.} Evolving legacy code is a challenging task, in particular when the code has been poorly written---
  which imposes significant cognitive load on developers when trying to familiarize themselves with a legacy code base. 
\noindent{\bf Problem.} 
%Besides lack of experience and other factors, 
Previous research have investigatedthe impact of small code patterns
that contribute to misunderstanding the source code written in \clang and \cpplang.
However, little is known about the influence of these small code patterns (hereafter atoms of confusion)
on understanding a dynamic languages (such as JavaScript). 
\noindent{\bf Research Method and Goal.} In this work, we present the results of a mixed-methods research:
a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language
whose popularity is growing in the most different application domains (from web programming to backend and IOT development).
\noindent{\bf Results.} From our survey and interviews, we confirm that atoms of confusion lead
to code that is hard to understand: 
developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not
present in 6 out of 9 atom candidates.
Also, the results of our MSR effort reveal that atoms of confusion frequently arise
in open-source JavaScript systems: the \texttt{comma operator} is present in 11\% of the projects in our dataset; while five other
atoms appear in more than 60\% of the projects---including the atoms Assignment as Value and Oomitted Curly Braces, which appear 8253 and 53357 times,
respectively. Our investigation also reveals additional
JavaScript constructs and idioms that might hinder developers from understanding
JavaScript code. For instance, even experienced JavaScript
developers agree that it is hard to understand code that uses callback functions,
even though we found widespread use of \texttt{callbacks} in our
projects dataset.
\noindent{\bf Implications.} Altogether, our findings might help
developers write cleaner JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code. 
\end{abstract}

>>>>

\begin{abstract}
Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase. Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++. However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript). In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort.
Our goal is to investigate the impact of atoms of confusion on understanding JavaScript code---a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development). From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates. Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates
appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, which appear in 8253 and 53357 times,
respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset. Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
\end{abstract}
